Bioverse: Exoplanet Biosignature Patterns Under the UV Threshold Hypothesis for the Origin of Life

### 2024-11-20 src/tex/ms.tex
update text based on power grid

Notably, the sensitivity of the M~dwarf sample extends into the low \gls{NUV} flux end due to the broader distribution of maximum past \gls{NUV} fluxes in this sample.
Here, the FGK sample is less sensitive.
\todo[inline]{Update this with final results from stat. power grid}
The only parameter space region with statistical power above \SI{90}{\percent} lies at abiogenesis rates $f_\mathrm{life} > 0.9$ and threshold \gls{NUV} fluxes of $\sim \SIrange{200}{400}{\erg\per\second\per\centi\meter\squared}$.

\begin{figure*}
    \script{powergrid.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/powergrid}
%        \includegraphics[width=\hsize]{figures/powergrid.bak.png}
        \caption{Statistical power as a function of threshold \gls{NUV} flux and abiogenesis rate. Even for a large sample (here: $N=\var{N_nautilus}$), a high statistical power of the transit survey requires high abiogenesis rates $f_\mathrm{life}$.
         Intermediate values of $F_\mathrm{NUV, min}$ are more likely to yield strong evidence than extreme values; the sensitivity of the M~dwarf sample extends into the low \gls{NUV} flux end.
         }
        \label{fig:powergrid}
    \end{centering}
\end{figure*}

@@

In our semi-analytical analysis (Section~\ref{sec:results-semianalytical}), we project a possible test performed by a future observer equipped with a sample of exoplanets with derived past maximum \gls{NUV} exposure for which biosignature detection has been attempted.
This is necessarily reductive as this observer will have more knowledge about experimental conditions and will therefore be able to use this information to guide hypothesis testing.
For instance, we have made the choice to consider the total number of detected biosignatures as our summary statistic (Equation~\ref{eq:semian:likelihoodH1}), which is not `sufficient' to infer $F_{\mathrm{NUV},min}$ and $f_{life}$ separately.
However, by conditioning the Bayes factor to these variables (Equation \ref{eq:bayes_factor}), we calculate the probability distribution of the Bayesian evidence in favor of $H_1$.
In doing so, we may evaluate how evidence depends on the uncertainty over these unknown parameters in general terms, without assuming which particular test a future observer might choose to actually perform over real data when available.
From this, we can see that target selection can strongly affect the conclusiveness of a future test of the UV~Threshold Hypothesis.

@@

%\subsection{Constraining power for the origins of life as a function of biosignature location} % corresponds to conclusion points 2, 5
\subsection{How planetary context may constrain the UV~Threshold Hypothesis} % corresponds to conclusion points 2, 5
It comes to no surprise that the success rate of such a test is sensitive to the sample size of the survey and to the occurrence of life on temperate exoplanets.
As we have shown, the statistical power of this test also depends on the distribution of past \gls{NUV} fluxes in the sample and on the threshold flux.
Optimizing the survey to sample a wide range of NUV flux values, particularly at the extremes, can enhance the likelihood of obtaining strong evidence for or against the hypothesis.
Intermediate values of the threshold \gls{NUV} flux are more likely to yield strong evidence than extreme values, as the dataset under the alternative hypothesis $H_1$ differs more from the null hypothesis in this case while still being sufficiently populated.
The threshold flux is, of course, a priori unknown and we cannot influence it.
If, however, better theoretical predictions for the required \gls{NUV} flux for abiogenesis become available~\citep{Rimmer2021}, the survey strategy can be further optimized, for instance by targeting planets that are estimated to have received a \gls{NUV} flux slightly below and above this threshold or by applying a bisection algorithm in a sequential survey~\citep{Fields2023}.

@@

On the other hand, M~dwarfs show a wider distribution of maximum past \gls{NUV} fluxes in their \glspl{HZ}.
While old M~dwarfs can be considered low-UV environments, a significant fraction of them emit high \gls{NUV} fluxes into their \gls{HZ} during their early stages, in particular later subtypes~\citep{Richey-Yowell2023}.
This will help to test the high \gls{NUV} flux end of the UV Threshold Hypothesis; a higher occurrence of biosignatures here would support the hypothesis that a higher NUV flux is beneficial or necessary for life.
At the same time, a fraction of host stars in our M~dwarf sample extends it to lower maximum past \gls{NUV} fluxes, enabling tests of the low \gls{NUV} flux end of the hypothesis.
The higher and more variable \gls{NUV} fluxes in M~dwarfs thus increase the likelihood of obtaining strong evidence for or against the UV~Threshold Hypothesis.
The combination of a lack of UV radiation today, which makes biosignature gases more detectable~\citep{Segura2005}, and a UV-rich past that may have enabled abiogenesis could make M~dwarfs the preferred targets for biosignature searches.

These findings underscore the importance of constraining the UV emission profiles of \gls{EEC} host stars throughout their evolutionary stages to assess the viability of M-dwarf planets as testbeds for theories on the origins of life~\citep{Rimmer2021,Ranjan2023a}.

@@

Our main findings are:
\begin{enumerate}
    \item The UV Threshold Hypothesis of the cyanosulfidic scenario for the origins of life should lead to a correlation between past \gls{NUV} flux and current occurrence of biosignatures that may be observationally testable.
    \item The required sample size for detecting this correlation depends on the abiogenesis rate on temperate exoplanets and the distribution of host star properties in the sample; in particular their maximum past \gls{NUV} fluxes.
    Samples smaller than 50 planets are unlikely to yield conclusive results.
    \item Under the UV~Threshold Hypothesis, the fraction of inhabited planets in a transit survey is sensitive to the threshold \gls{NUV} flux and is expected to drop sharply for required fluxes above a few hundred~\SI{}{\erg\per\second\per\centi\meter\squared}.
    \item If the predicted UV~correlation exists, obtaining strong evidence for the hypothesis is likely ($\gtrsim \SI{80}{\percent}$) for sample sizes $\geq 100$ if the abiogenesis rate is high ($\gtrsim \SI{50}{\percent}$) and if no very high \gls{NUV} fluxes are required.
    A survey strategy that targets extreme values of inferred past \gls{NUV} irradiation increases the diagnostic power.
    \item Samples of planets orbiting M~dwarfs overall yield higher chances of successfully testing the UV~Threshold Hypothesis.
          They may also be more likely to yield biosignature detections under this hypothesis.
          \todo[inline]{check if new grid still supports this}

>>>>

Parameter space regions with statistical power above \SI{90}{\percent} lie at abiogenesis rates $f_\mathrm{life} > 0.5$ and mostly at threshold \gls{NUV} fluxes of $\sim \SIrange{200}{400}{\erg\per\second\per\centi\meter\squared}$.
Notably, the sensitivity of the M~dwarf sample extends into the low \gls{NUV} flux end due to the broader distribution of maximum past \gls{NUV} fluxes in this sample.
Here, the FGK sample is barely sensitive.

\begin{figure*}
    \script{powergrid.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/powergrid}
        \caption{Statistical power as a function of threshold \gls{NUV} flux and abiogenesis rate. Even for a large sample (here: $N=\var{N_nautilus}$), a high statistical power of the transit survey requires high abiogenesis rates $f_\mathrm{life}$.
         Intermediate values of $F_\mathrm{NUV, min}$ are more likely to yield strong evidence than extreme values. For abiogenesis rates $f_\mathrm{life} \gtrsim \SI{80}{\percent}$, the sensitivity of the M~dwarf sample extends into the low \gls{NUV} flux end.
         }
        \label{fig:powergrid}
    \end{centering}
\end{figure*}

@@

In our semi-analytical analysis (Section~\ref{sec:results-semianalytical}), we project a possible test performed by a future observer equipped with a sample of exoplanets with derived past maximum \gls{NUV} exposure for which biosignature detection has been attempted.
This is necessarily reductive as this observer will have more knowledge about experimental conditions and will therefore be able to use this information to guide hypothesis testing.
For instance, we have made the choice to consider the total number of detected biosignatures as our summary statistic (Equation~\ref{eq:semian:likelihoodH1}), which is not sufficient to infer $F_{\mathrm{NUV},min}$ and $f_{life}$ separately.
However, by conditioning the Bayes factor to these variables (Equation \ref{eq:bayes_factor}), we calculate the probability distribution of the Bayesian evidence in favor of $H_1$.
In doing so, we may evaluate how evidence depends on the uncertainty over these unknown parameters in general terms, without assuming which particular test a future observer might choose to actually perform over real data when available.
From this, we can see that target selection can strongly affect the conclusiveness of a future test of the UV~Threshold Hypothesis.

@@

%\subsection{Constraining power for the origins of life as a function of biosignature location} % corresponds to conclusion points 2, 5
\subsection{How planetary context may constrain the UV~Threshold Hypothesis} % corresponds to conclusion points 2, 5
It comes to no surprise that the success rate for testing the UV~Threshold Hypothesis is sensitive to the sample size of the survey and to the occurrence of life on temperate exoplanets.
As we have shown, the statistical power of this test also depends on the distribution of past \gls{NUV} fluxes in the sample and on the threshold flux.
Optimizing the survey to sample a wide range of NUV flux values, particularly at the extremes, can enhance the likelihood of obtaining strong evidence for or against the hypothesis.
Intermediate values of the threshold \gls{NUV} flux are more likely to yield strong evidence than extreme values, as the dataset under the alternative hypothesis $H_1$ differs more from the null hypothesis in this case while still being sufficiently populated.
The threshold flux is, of course, a priori unknown and we cannot influence it.
If, however, better theoretical predictions for the required \gls{NUV} flux for abiogenesis become available~\citep{Rimmer2021}, the survey strategy can be further optimized, for instance by targeting planets that are estimated to have received a \gls{NUV} flux slightly below and above this threshold or by applying a bisection algorithm in a sequential survey~\citep{Fields2023}.

@@

On the other hand, M~dwarfs show a wider distribution of maximum past \gls{NUV} fluxes in their \glspl{HZ}.
While old M~dwarfs can be considered low-UV environments, a significant fraction of them emit high \gls{NUV} fluxes into their \gls{HZ} during their early stages, in particular later subtypes~\citep{Richey-Yowell2023}.
This will help to test the high \gls{NUV} flux end of the UV Threshold Hypothesis; a higher occurrence of biosignatures here would support the hypothesis that a higher NUV flux is beneficial or necessary for life.
At the same time, a fraction of host stars in our M~dwarf sample extends it to lower maximum past \gls{NUV} fluxes, enabling tests of the low \gls{NUV} flux end of the hypothesis.
The higher and more variable \gls{NUV} fluxes in M~dwarfs thus increase the likelihood of obtaining strong evidence for or against the UV~Threshold Hypothesis.
The combination of a lack of UV radiation today, which makes biosignature gases more detectable~\citep{Segura2005}, and a UV-rich past that may have enabled abiogenesis could make M~dwarfs the preferred targets for biosignature searches.
These findings underscore the importance of constraining the UV emission profiles of \gls{EEC} host stars throughout their evolutionary stages to assess the viability of M-dwarf planets as testbeds for theories on the origins of life~\citep{Rimmer2021,Ranjan2023a}.

@@

Our main findings are:
\begin{enumerate}
    \item The UV Threshold Hypothesis of the cyanosulfidic scenario for the origins of life should lead to a correlation between past \gls{NUV} flux and current occurrence of biosignatures that may be observationally testable.
    \item The required sample size for detecting this correlation depends on the abiogenesis rate on temperate exoplanets and the distribution of host star properties in the sample; in particular their maximum past \gls{NUV} fluxes.
    Samples smaller than 50 planets are unlikely to yield conclusive results.
    \item Under the UV~Threshold Hypothesis, the fraction of inhabited planets in a transit survey is sensitive to the threshold \gls{NUV} flux and is expected to drop sharply for required fluxes above a few hundred~\SI{}{\erg\per\second\per\centi\meter\squared}.
    \item If the predicted UV~correlation exists, obtaining strong evidence for the hypothesis is likely ($\gtrsim \SI{80}{\percent}$) for sample sizes $\geq 100$ if the abiogenesis rate is high ($\gtrsim \SI{50}{\percent}$) and if no very high \gls{NUV} fluxes are required.
    A survey strategy that targets extreme values of inferred past \gls{NUV} irradiation increases the diagnostic power.
    \item Samples of planets orbiting M~dwarfs overall yield higher chances of successfully testing the UV~Threshold Hypothesis.
          They may also be more likely to yield biosignature detections under this hypothesis.
